Appendix: Why the complement of AUC isn't a casewise CV criterion

### 2023-12-31 scratch/notes_georges/AUC_not_casewise-edited-1.Rmd
show dimensions of matrices in AUC appendix

Consider calculating AUC for folds in which a validation set contains $n_v$ observations. To calculate AUC in the validation set, we need the vector of prediction criteria, $\widehat{\mathbf{y}} = (\widehat{y}_1, ..., \widehat{y}_{n_v})^T$, and the vector of observed responses in the validation set, $\mathbf{y} = (y_1, \ldots, y_{n_v})^T$ with $y_i \in \{0,1\}, \; i = 1, \ldots, n_v$.

@@

We then define a function to generate all possible $\mathbf{y}$s of length $n_v$ as rows of the matrix $\mathbf{Y}$:

@@

The values of $\mathrm{cv}(\widehat{y}_i, y_i)$ that express AUC as a sum of casewise values are solutions of equation (\ref{eq:cw}), which can be written as solutions of the following linear equation:
\begin{equation}
\label{eq:lin}
\tag{2}
(\mathbf{U} -\mathbf{Y}) \mathbf{c}_0 + \mathbf{Y} \mathbf{c}_1
=
[\mathbf{U} -\mathbf{Y}, \mathbf{Y}]
\begin{bmatrix}
\mathbf{c}_0 \\ \mathbf{c}_1
\end{bmatrix}
= \mathrm{AUC}(\mathbf{\widehat{Y}},\mathbf{Y})
\end{equation}
where $\mathbf{U}$ is a matrix of 1s conformable with $\mathbf{Y}$; $\mathbf{c}_0 = [\mathrm{cv}(1,0), c(2,0), ..., \mathrm{cv}(n_v,0)]^T$; $\mathbf{c}_1 = [\mathrm{cv}(1,1), c(2,1), ..., \mathrm{cv}(n_v,1)]^T$; $[\mathbf{U} -\mathbf{Y}, \mathbf{Y}]$ and
$\begin{bmatrix}\begin{aligned}
\mathbf{c}_0 \\ \mathbf{c}_1
\end{aligned}
\end{bmatrix}$ 
are partitioned matrices; and $\mathbf{\widehat{Y}}$ is a matrix each of whose rows consists of the integers 1 to $n_v$.

We can test whether equation (\ref{eq:lin}) has a solution for any given $n_v$ by trying to solve it as a least-squares problem, considering whether the residuals of the associated linear model are all 0, using the "design matrix" $[\mathbf{U} -\mathbf{Y}, \mathbf{Y}]$ to predict the "outcome"
$\mathrm{AUC}(\mathbf{\widehat{Y}},\mathbf{Y})$:

>>>>

Consider calculating AUC for folds in which a validation set contains $n_v$ observations. To calculate AUC in the validation set, we need the vector of prediction criteria, $\widehat{\mathbf{y}}_{(n_v \times 1)} = (\widehat{y}_1, ..., \widehat{y}_{n_v})^T$, and the vector of observed responses in the validation set, $\mathbf{y}_{(n_v \times 1)} = (y_1, \ldots, y_{n_v})^T$ with $y_i \in \{0,1\}, \; i = 1, \ldots, n_v$.

@@

We then define a function to generate all possible $\mathbf{y}$s of length $n_v$ as rows of the matrix $\mathbf{Y}_{(2^{n_v} \times n_v)}$:

@@

The values of $\mathrm{cv}(\widehat{y}_i, y_i)$ that express AUC as a sum of casewise values are solutions of equation (\ref{eq:cw}), which can be written as solutions of the following linear equation:
\begin{equation}
\label{eq:lin}
\tag{2}
(\mathbf{U} -\mathbf{Y}) \mathbf{c}_0 + \mathbf{Y} \mathbf{c}_1
=
[\mathbf{U} -\mathbf{Y}, \mathbf{Y}]
\begin{bmatrix}
\mathbf{c}_0 \\ \mathbf{c}_1
\end{bmatrix}
= \mathrm{AUC}(\mathbf{\widehat{Y}},\mathbf{Y})
\end{equation}
where $\mathbf{U}_{(2^{n_v} \times n_v)}$ is a matrix of 1s conformable with $\mathbf{Y}$; $\mathbf{c}_0 = [\mathrm{cv}(1,0), c(2,0), ..., \mathrm{cv}(n_v,0)]^T$; $\mathbf{c}_1 = [\mathrm{cv}(1,1), c(2,1), ..., \mathrm{cv}(n_v,1)]^T$; $[\mathbf{U} -\mathbf{Y}, \mathbf{Y}]_{(2^{n_v} \times 2n_v)}$ and
$\begin{bmatrix}\begin{aligned}
\mathbf{c}_0 \\ \mathbf{c}_1
\end{aligned}
\end{bmatrix}_{(2n_v \times 1)}$ 
are partitioned matrices; and $\mathbf{\widehat{Y}}_{(2^{n_v} \times n_v)}$ is a matrix each of whose rows consists of the integers 1 to $n_v$.

We can test whether equation (\ref{eq:lin}) has a solution for any given $n_v$ by trying to solve it as a least-squares problem, considering whether the residuals of the associated linear model are all 0, using the "design matrix" $[\mathbf{U} -\mathbf{Y}, \mathbf{Y}]$ to predict the "outcome"
$\mathrm{AUC}(\mathbf{\widehat{Y}},\mathbf{Y})_{(2n_v \times 1)}$:
